---
title: "Bayesian Estimation with BetaDanish"
author: "Bilal Ahmad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Bayesian Estimation with BetaDanish}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE
)
have_mcmc <- requireNamespace("MCMCpack", quietly = TRUE) &&
             requireNamespace("coda",     quietly = TRUE)
knitr::opts_chunk$set(eval = have_mcmc)
```

## 1. Overview

This vignette walks through Bayesian estimation of the three-parameter
Exponentiated Danish (ED) submodel and the full four-parameter
Beta-Danish distribution using `bayes_betadanish()`. The sampler is a
random-walk Metropolis run on log-transformed parameters with vague
Gamma(0.01, 0.01) priors, and results are returned on the natural
scale.

`MCMCpack` and `coda` are Suggests rather than hard dependencies; this
vignette is evaluated only when both are installed.

## 2. ED submodel on bladder cancer remission times

```{r}
library(BetaDanish)
data("remission")
fit_bayes <- bayes_betadanish(
  time     = remission$time,
  status   = remission$status,
  submodel = TRUE,
  burnin   = 2000, mcmc = 5000,
  tune     = 0.5,
  seed     = 1
)
print(fit_bayes)
```

## 3. Posterior diagnostics

```{r}
draws <- fit_bayes$draws
op <- par(mfrow = c(3, 1), mar = c(3, 4, 2, 1))
coda::traceplot(draws)
par(op)
op <- par(mfrow = c(3, 1), mar = c(3, 4, 2, 1))
coda::autocorr.plot(draws, auto.layout = FALSE)
par(op)
```

## 4. Posterior survival vs Kaplan-Meier

```{r}
post_mean <- summary(draws)$statistics[, "Mean"]
b <- post_mean["b"]; c <- post_mean["c"]; k <- post_mean["k"]
km <- survival::survfit(survival::Surv(time, status) ~ 1, data = remission)
plot(km, conf.int = FALSE, xlab = "Time (months)",
     ylab = "Survival probability",
     main = "Posterior mean ED fit on remission data")
t_grid <- seq(0.1, max(remission$time), length.out = 200)
S_post <- pbetadanish(t_grid, a = 1, b = b, c = c, k = k,
                      lower.tail = FALSE)
lines(t_grid, S_post, col = "red", lwd = 2)
legend("topright",
       legend = c("Kaplan-Meier", "Posterior-mean ED"),
       col = c("black", "red"), lty = 1, lwd = c(1, 2), bty = "n")
```

## 5. Bayesian vs MLE comparison

```{r}
fit_mle <- fit_betadanish(survival::Surv(time, status) ~ 1,
                          data = remission, submodel = TRUE,
                          n_starts = 3)
cbind(
  MLE  = round(coef(fit_mle), 4),
  Bayes_post_mean = round(post_mean, 4)
)
```

The posterior means and MLE point estimates are typically very close
for this dataset, with the Bayesian credible intervals offering a more
honest accounting of estimation uncertainty than asymptotic Wald
intervals.
