---
title: "Cure Models with the Beta-Danish Distribution"
author: "Bilal Ahmad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cure Models with the Beta-Danish Distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE
)
```

## 1. Two cure formulations

When a non-trivial fraction of subjects never experience the event,
standard parametric survival models are misspecified. The package
implements two complementary cure formulations on the Exponentiated
Danish kernel:

- The **mixture** cure model splits the population into a susceptible
  fraction \(\pi(Z)\) (modelled by logistic regression on incidence
  covariates) and a cured fraction \(1 - \pi(Z)\).
- The **promotion-time** (non-mixture) cure model derives the cure
  fraction from a latent Poisson process of clonogenic cells with
  intensity \(\theta(Z) = \exp(Z' \gamma)\); the cure fraction is
  \(\exp(-\theta(Z))\).

Both are fitted by `fit_bd_cure(type = "mixture")` and
`type = "promotion"`, sharing the same AFT latency component
\(k_i = \exp(X_i' \delta)\).

## 2. Bone marrow transplant data

```{r}
library(BetaDanish)
data("transplant")
str(transplant)
```

The `transplant` dataset contains 91 patients with a treatment group
indicator and a known cure plateau in the Kaplan-Meier estimator.

## 3. Promotion-time cure model

```{r}
fit_prom <- fit_bd_cure(
  formula_aft  = survival::Surv(time, status) ~ 1,
  formula_cure = ~ group,
  data         = transplant,
  type         = "promotion",
  n_starts     = 3
)
summary(fit_prom)
```

## 4. Mixture cure model

```{r}
fit_mix <- fit_bd_cure(
  formula_aft  = survival::Surv(time, status) ~ 1,
  formula_cure = ~ group,
  data         = transplant,
  type         = "mixture",
  n_starts     = 3
)
summary(fit_mix)
```

## 5. Model comparison

```{r}
data.frame(
  Model  = c("Promotion-time", "Mixture"),
  logLik = c(fit_prom$logLik, fit_mix$logLik),
  AIC    = c(fit_prom$AIC,    fit_mix$AIC),
  BIC    = c(fit_prom$BIC,    fit_mix$BIC)
)
```

## 6. Cox-Snell residual diagnostic

```{r}
plot(fit_mix)
```

A diagonal-aligned plot indicates the cure-model specification is
adequate; substantial departure from \(y = x\) suggests the latency
distribution or the link function for the cure component need
revisiting.
