---
title: "Competing Risks with the Beta-Danish Distribution"
author: "Bilal Ahmad"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Competing Risks with the Beta-Danish Distribution}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 4.5,
  warning = FALSE,
  message = FALSE
)
have_cmprsk <- requireNamespace("cmprsk", quietly = TRUE)
knitr::opts_chunk$set(eval = have_cmprsk)
```

## 1. Overview

Under independent latent failure times the joint competing-risks
likelihood factorises into one Beta-Danish marginal per cause. The
`fit_bd_competing()` function fits all causes jointly via
bound-constrained multi-start L-BFGS-B, and `cif_compare()` overlays the
fitted cumulative incidence functions on the nonparametric Aalen-Johansen
estimator and runs Gray's test.

`cmprsk` is a Suggests; this vignette is evaluated only when installed.

## 2. Simulated two-cause data

```{r}
library(BetaDanish)
set.seed(2026)
n <- 400
T1 <- rbetadanish(n, a = 1.2, b = 1.5, c = 1.0, k = 0.4)
T2 <- rbetadanish(n, a = 1.0, b = 2.0, c = 1.0, k = 0.2)
C  <- stats::rexp(n, 0.05)
time  <- pmin(T1, T2, C)
cause <- ifelse(time == C, 0L,
                ifelse(T1 <= T2, 1L, 2L))
table(cause)
```

## 3. Fitting the joint competing-risks model

```{r}
fit <- fit_bd_competing(time = time, cause = cause,
                        n_restarts = 5, seed = 1)
print(fit)
summary(fit)
```

## 4. Cause-specific cumulative incidence functions

```{r}
res <- cif_compare(fit, plot = TRUE)
```

The faceted overlay places the model-based CIF against the
Aalen-Johansen step function for each cause; close agreement supports
the independent-latent-times specification.

## 5. Gray's test

```{r, eval = !is.null(res$gray_test)}
res$gray_test
```

When Gray's test rejects equality of the CIFs across causes, the
biological interpretation is that the two latent failure mechanisms have
materially different long-run probabilities of being the cause of
failure.

## 6. Closing remarks

The pattern shown here generalises to three or more causes; the
`fit_bd_competing()` interface accepts any `0, 1, ..., m` cause coding
where `0` indicates right-censoring. Real-data illustrations on the
`bmt`, `follic`, and `mgus2` datasets follow the same workflow.
